---
title: "readme"
#author: "Paul Q. Sims"
#date: "7/13/2020"
output: 
  github_document
---

## Description

Data, code, and figures associated with the study "Repeatable differences in snowshoe hare behavior: Animal personality may facilitate adaption to climate change". *In prep*. Please see the publication for further details on data collection, the experimental apparatus, procedure, and statistical analyses.

This repository will be permanently archived at Zenodo (link to be provided), with the latest release to be found here: <!--[![DOI](https://zenodo.org/badge/279408355.svg)](https://zenodo.org/badge/latestdoi/279408355).-->

A report summary containing all the R script output, including plots, can be found at [analysis_summary.md](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/reports/analysis_summary.md) in `reports/`. Markdown file versions (.md) are recommended for viewing on github and HMTL file versions (.html) are recommended for desktop viewing.

R Markdown files for the readme and summary report can be rendered by sourcing [render-Rmd.R](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/R/render-Rmd.R) in `R/`. Individual .Rmd files for each section of the analysis can be rendered using the provided code in the `render-Rmd.R` file or by knitting each .Rmd file independently.

**Last update**: `r Sys.Date()`

## Authors

Diana J.R. Lafferty (<dlaffert@nmu.edu>)

Paul Q. Sims (<paul.q.sims@gmail.com>)

L. Scott Mills (<scott.mills@mso.umt.edu>)

## Directory structure

* `data/` Data sets used in the analyses
* `R/` Self-contained R markdown (*.Rmd) scripts (can be run on their own) and custom functions used for modifying the data, running analyses, and generating plots
* `figs/` High quality figures generated for publication
* `reports/` Markdown and html summaries of code output

## Prerequisites: R version and R packages

`r R.Version()$version.string`

```{r package names, echo=FALSE,warning=FALSE,eval=TRUE,message=FALSE}
# Package dataframe
dat_packages <- setNames(data.frame(matrix(ncol = 3, nrow = 15)),
                        c("Package_Name", "Version", "Function"))
# Package name
dat_packages$Package_Name <- 
  c("tidyverse", "broom.mixed", "broom",
    "glmmTMB", "MuMIn", "lme4", "lmerTest", "DHARMa",
    "ggplot2", "ggsignif", "scales", "effects",
    "rprojroot", "rmarkdown", "knitr")

# Package function
dat_packages$Function <- 
  c(rep("Data clean up", times = 3), 
    rep("Analyses", time = 5),
    rep("Plots", times = 4),
    rep("Miscellaneous", times = 3))
# Package version
for (i in 1:nrow(dat_packages)) {
   dat_packages$Version[i] <- as.character(packageVersion(dat_packages$Package_Name[i]))
}
knitr::kable(dat_packages)
```

## Metadata

Reference files: 

- [data_activity_LaffertyEtAl_2020.csv](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/data/data_activity_LaffertyEtAl_2020.csv)
- [data_day_LaffertyEtAl_2020.csv](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/data/data_day_LaffertyEtAl_2020.csv)
- [data_time_LaffertyEtAl_2020.csv](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/data/data_time_LaffertyEtAl_2020.csv)

```{r, echo = FALSE, results = TRUE, message = FALSE}

library(tidyverse)

var_names_act <- 
  read_csv("data/data_activity_LaffertyEtAl_2020.csv") %>%
  names(.)
var_names_time <- 
  read_csv("data/data_time_LaffertyEtAl_2020.csv") %>%
  names(.)
var_names_day <- 
  read_csv("data/data_day_LaffertyEtAl_2020.csv") %>%
  names(.)

# Write out descriptions associated with each column header
var_names <- 
  c(var_names_act, var_names_time, var_names_day) %>%
  as_tibble(.) %>%
  distinct(.) 

var_descrip <- c(
  "Individual ID",
  "Trial: 1 or 2",
  "Date of measurement",
  "Time point: 1-23",
  "Background color for activity data: Light or dark",
  "Open-Field location for activity data: Inner or outer",
  "Lines crossed",
  "Time spent in the shelter",
  "Sex: Female or male",
  "Mass in g",
  "Mass scaled by 1 standard deviation and mean centered",
  "ID for each individual by trial",
  "Time point scaled by 1 standard deviation and mean centered",
  "Background color for non-activity data: Light or dark",
  "Open-Field location for non-activity data: Inner or outer",
  "Shelter status: Inside or outside",
  "Time (s) to exit the shelter",
  "Whether or not hares exited the shelter immediately upon release: Yes or no",
  "Number of unique tiles entered over the course of the day",
  "Total number of unique tiles available in the open-field",
  "Number of unique tiles hare did not enter out of the total")

var_metadata <- tibble(var_names,var_descrip) %>%
                  rename(Column = "value", Description = "var_descrip")

knitr::kable(var_metadata)
```

- Missing values are input as `NA` text

## Licenses

Software and code is licensed under the MIT License, data is licensed under CC0, and media files are under CC-BY-4.0 - see the [license.md](https://github.com/paulqsims/snowshoe-hare_personality/blob/master/license.md) file for details.

## Funding information

This work was supported by North Carolina State University, University of Montana, and the National Science Foundation Division of Environmental Biology (grant #1354449 to L.S.M.).

