# Repeatability analyses
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Repeatability analyses and variance estimates for Lafferty et al. 2020

```{r rpt-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE, cache = TRUE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE, cache = TRUE)
```

## Setup

```{r rpt-setup, cache=FALSE}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(lme4)
library(lmerTest)
library(DHARMa)
library(broom.mixed)  # for tidying model output
library(knitr)
library(glmmTMB)  # for activity analyses

# Load personalized functions
source("R/custom-functions.R")

# Read in data 
# Data collected once each day
data_day <-
  read_csv("data/data_day_LaffertyEtAl_2020.csv",
           col_types = list(exit_fast = col_factor(),
                            trial = col_factor(levels = c("1", "2")),
                            sex = col_factor(levels = c("Female", "Male")))) %>%
  mutate(obsID = 1:nrow(.))  # observation level random effect


# Data collected at time points throughout the day
data_time <-
  read_csv("data/data_time_LaffertyEtAl_2020.csv",
             col_types = list(trial = col_factor(levels = c("1", "2")),
                              sex = col_factor(levels = c("Female", "Male")))) %>%
  mutate(obsID = 1:nrow(.),  # observation level random effect
         across(where(is.character), ~ as_factor(.x)))  

# Data collected at time points throughout the day for lines crossed
data_act <-
  read_csv("data/data_activity_LaffertyEtAl_2020.csv",
             col_types = list(trial = col_factor(levels = c("1", "2")),
                              sex = col_factor(levels = c("Female", "Male")))) %>%
  mutate(obsID = 1:nrow(.),  # observation level random effect
         across(where(is.character), ~ as_factor(.x)))  
```

## Area explored

- Proportion of unique tiles entered

**Create dataset without NAs**

```{r rpt-expl-dat}
data_expl_rpt <- 
  remove_dat_na(data_day, c("uniq_tiles", "tile_fails", "trial",
                            "sex", "ID", "day", "obsID")) %>%
  mutate(expl.y = cbind(uniq_tiles, tile_fails))
```


### Random effect selection

**Test significance of individual intercept (1|ID)**

```{r rpt-expl-ID}
# Fit model for adjusted repeatability
m_AdjRpt_expl <- 
  glmer(expl.y ~ trial + sex + (1|ID),  
        family = "binomial", 
        data = data_expl_rpt,
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Fit null model for testing significance of individual ID
m_AdjRpt_expl_null <- 
  glm(expl.y ~ trial + sex,  
        family = "binomial", 
        data = data_expl_rpt,
        contrasts = list(trial = c(-1,1), sex = c(-1,1)))

# Likelihood ratio test for significance of ind. random effect and rpt
#  Custom functions
rpt_SigTest(m_AdjRpt_expl, m_AdjRpt_expl_null)
```

### Variance and repeatability estimates

```{r expl-var-rpt-est, cache=TRUE}
# Adjusted repeatability and variance estimates
# Bootstrap 95% C.I. 
res_boot <- bootMer(m_AdjRpt_expl,  # final rpt model
                    use.u = FALSE,  
                    FUN = rpt_day,  # custom rpt function
                    seed = 1,  # allows exact replication of results
                    nsim = 1000,
                    type = "parametric",
                    parallel = "multicore",
                    .progress = "txt")

# Calculate 95% CI
RptConfInt_day(res_boot)
```

## Risk-taking

### Shelter exit latency

- Exit immediately: yes or no

**Create data set with NA values removed**

```{r shel-exit-data}
data_shelExit_rpt <- 
  remove_dat_na(data_day, c("exit_fast", "trial",
                            "sex", "ID", "day",
                            "obsID"))
```

#### Random effect selection

**Test significance of individual intercept (1|ID)**

```{r shel-exit-ID, cache=TRUE}
# Fit model for adjusted repeatability
m_AdjRpt_ExitLat <- 
  glmer(exit_fast ~ trial + sex + (1|ID),  
        family = "binomial", 
        data = data_shelExit_rpt,
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Fit null model for testing significance of individual ID
m_AdjRpt_ExitLat_null <- 
  glm(exit_fast ~ trial + sex,  
        family = "binomial", 
        data = data_shelExit_rpt,
        contrasts = list(trial = c(-1,1), sex = c(-1,1)))

# Likelihood ratio test for significance of ind. random effect and rpt
rpt_SigTest(m_AdjRpt_ExitLat, m_AdjRpt_ExitLat_null)
```

#### Variance and repeatability estimates

```{r shel-exit-var-rpt-est}
# Variance estimates, including Adjusted repeatability 
rpt_day(m_AdjRpt_ExitLat) %>%
  pretty_DayRpt(.)
```

- 95% C.I. not computed since individual ID is non-significant

### Shelter preference

- Inside or outside the shelter

**Create data set with NA values removed**

```{r shel-data}
data_shel_rpt <- 
  remove_dat_na(data_time, c("shelter", "t_point_sc", "trial",
                             "sex", "ID", "trial_series", "day",
                             "obsID"))
```

#### Random-effect selection

**Test significance of individual intercept (1|ID)**

```{r shel-ID}
# Fit base model
m_shel_base <- 
  glmer(shelter ~ t_point_sc * trial + sex +
          (1|ID) + (t_point_sc|trial_series) + (1|obsID),
        data = data_shel_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Fit model without individual ID
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(t_point_sc|trial_series) + (1|obsID)")

# LRT for significance of individual ID
rpt_SigTest(m_shel_base, m_shel_reduc)

# Non-significant random intercept variance estimate
kable(as_tibble_col(VarCorr(m_shel_base)$ID[1],
      column_name = "Random intercept variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of individual intercept and time slope correlation (t_point_sc|ID)**

- Run before testing random slope because result of slope intercept will determine which slope model to use

```{r shel-ID-time-corr}
# Fit model with intercept-slope correlation
m_shel_full <- refitModelRanEf(m_shel_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series) +
                                  (1|obsID)")  

# Fit model without intercept-slope m_shel_base
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series) + (1|obsID)")

# LRT for significance of intercept-slope correlation
rpt_SigTest(m_shel_full, m_shel_reduc)

# print non-significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_shel_full)$ID,"correlation")[2,1],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2,
      align = "l")
```

- Boundary fit due to overly complex random effect structure, resulting in individual id estimated as zero and random intercept-slope correlation as NaN

**Test significance of time slope (t_point_sc)**

```{r shel-time}
# Fit model with random slope
m_shel_full <- refitModelRanEf(m_shel_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series) + (1|obsID)")  

# Fit model without random slope
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (t_point_sc|trial_series) + (1|obsID)")

# LRT for significance of random slope 
rpt_SigTest(m_shel_full, m_shel_reduc)

# Non-significant random slope variance estimate
kable(as_tibble_col(VarCorr(m_shel_full)$ID.1[1],
      column_name = "Random slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of (reaction norm) trial series intercept**

```{r shel-rxn-norm-trial-series-int}
# Fit model with random intercept
m_shel_full <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (1|trial_series)")

# Fit model without random intercept
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(1|ID)")

# LRT for significance of intercept
rpt_SigTest(m_shel_full, m_shel_reduc)
```

**Test significance of (reaction norm) trial series intercept and time slope correlation**

- Run before testing random slope because result of slope intercept will determine which slope model to use

```{r shel-rxn-norm-trial-series-time-slope-corr}
# Fit model with random intercept-slope correlation
m_shel_full <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (t_point_sc|trial_series) + (1|obsID)")  

# Fit model without random intercept-slope correlation
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (0 + t_point_sc|trial_series) +
                                  (1|trial_series) + (1|obsID)")

# LRT for significance of intercept-slope correlation
rpt_SigTest(m_shel_full, m_shel_reduc)
```

**Test significance of (reaction norm) time slope**

```{r shel-rxn-norm-time-slope}
# Fit model with random slope
m_shel_full <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (t_point_sc|trial_series) + (1|obsID)")  

# Fit model without random slope
m_shel_reduc <- refitModelRanEf(m_shel_base,
                                 "(1|ID) + (1|trial_series) + (1|obsID)")

# LRT for significance of random slope
rpt_SigTest(m_shel_full, m_shel_reduc)
```

**Final random effect model structure**

- Removed trial series intercept and time slope correlation due to model convergence errors during bootstrapping

```{r shel-final-rpt-mod}
m_shel_rpt_final <- 
  glmer(shelter ~ t_point_sc * trial + sex +
          (0 + t_point_sc|trial_series) + (1|trial_series),  
        data = data_shel_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))
```

#### Variance estimates

- Repeatability not calculated since individual intercept and time slope estimates were non-significant
- Removed random series intercept-slope covariance due to convergence errors

```{r shel-var-rpt-est, cache=TRUE}
# Bootstrap confidence intervals
res_bootOut <-
  bootMer(m_shel_rpt_final,  # final rpt model
          use.u = FALSE,
          FUN = varEst_shel,  # custom variance estimates function
          seed = 1,
          nsim = 1000,
          type = "parametric",
          parallel = "multicore",
          .progress = "txt")

# Compile variance estimates 
res_varEst <-
  as_tibble(res_bootOut$t) %>%
  rename("Series Intercept" = "V1",
         "Series Slope" = "V2",
         "Residual" = "V3") %>%
  map_df(~ quantile(.x, c((1 - 0.95)/2, 1 - (1 - 0.95)/2),
                  na.rm = TRUE), .id = "scale") %>%
  rename(Variable = scale, Lower95CI = `2.5%`, Upper95CI = `97.5%`) %>%
  mutate(Estimate = pluck(res_bootOut$t0)) %>%  # add R estimates
  relocate(Variable, Estimate) %>%  # reorder columns
  mutate_at(c(2:4), round, 2)  # # round estimates and errors to 2 digits

kable(res_varEst)
```

### Open-Field location preference

- Inner vs. outer area of the Open-Field

**Create data set with NA values removed**

```{r OF-data}
data_of_rpt <- 
  remove_dat_na(data_time, c("of_loc", "t_point_sc", "trial",
                             "sex", "ID", "trial_series", "day",
                             "obsID"))
```

#### Random-effect selection

**Test significance of individual intercept (1|ID)**

```{r OF-ID}
# Fit base model
m_OFloc_base <- 
  glmer(of_loc ~ t_point_sc * trial + sex +
          (1|ID) + (t_point_sc|trial_series),  
        data = data_of_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Fit model without individual ID
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(t_point_sc|trial_series)")

# LRT for significance of individual ID
rpt_SigTest(m_OFloc_base, m_OFloc_reduc)
```

**Test significance of individual intercept and time slope correlation (t_point_sc|ID)**

- Run before testing random slope because result of slope intercept will determine which slope model to use

```{r OF-ID-time-corr}
# Fit model with intercept-slope correlation
m_OFloc_full <- refitModelRanEf(m_OFloc_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series)")  

# Fit model without intercept-slope m_OFloc_base
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series)")

# LRT for significance of intercept-slope correlation
rpt_SigTest(m_OFloc_full, m_OFloc_reduc)

# print non-significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_OFloc_full)$ID,"correlation")[2,1],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of time slope (t_point_sc)**

```{r OF-time}
# Fit model with random slope
m_OFloc_full <- refitModelRanEf(m_OFloc_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series)") 

# Fit model without random slope
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (t_point_sc|trial_series)")

# LRT for significance of random slope 
rpt_SigTest(m_OFloc_full, m_OFloc_reduc)

# Non-significant random slope variance estimate
kable(as_tibble_col(VarCorr(m_OFloc_full)$ID.1[1],
      column_name = "Random slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of (reaction norm) trial series intercept**

```{r OF-rxn-norm-trial-series-int}
# Fit model with random intercept
m_OFloc_full <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (1|trial_series)")  

# Fit model without random intercept
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID)")

# LRT for significance of intercept
rpt_SigTest(m_OFloc_full, m_OFloc_reduc)
```

**Test significance of (reaction norm) trial series intercept and time slope correlation**

- Run before testing random slope b/c result of slope intercept will determine which slope model to use

```{r OF-rxn-norm-trial-series-time-slope-corr}
# Fit model with random intercept-slope correlation
m_OFloc_full <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (t_point_sc|trial_series)")  

# Fit model without random intercept-slope correlation
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (0 + t_point_sc|trial_series) +
                                  (1|trial_series)")

# LRT for significance of intercept-slope correlation
rpt_SigTest(m_OFloc_full, m_OFloc_reduc)
```

**Test significance of (reaction norm) time slope**

```{r OF-rxn-norm-time-slope}
# Fit model with random slope
m_OFloc_full <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (t_point_sc|trial_series)")  

# Fit model without random slope
m_OFloc_reduc <- refitModelRanEf(m_OFloc_base,
                                 "(1|ID) + (1|trial_series)")

# LRT for significance of random slope
rpt_SigTest(m_OFloc_full, m_OFloc_reduc)
```

**Final random effect model structure**

- Removed trial series intercept and time slope correlation due to model convergence errors during bootstrapping

```{r OF-final-rpt-mod}
m_OFloc_rpt_final <- 
  glmer(of_loc ~ t_point_sc * trial + sex +
          (1|ID) + (0 + t_point_sc|trial_series) + (1|trial_series),  
        data = data_of_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))
```

#### Variance and repeatability estimates

```{r OF-var-rpt-est, cache=TRUE}
# Bootstrap confidence intervals
res_bootOut <- 
  bootMer(m_OFloc_rpt_final,  # final rpt model 
          use.u = FALSE, 
          FUN = varEst_OF,  # custom variance estimates function
          seed = 1,
          nsim = 1000,  
          type = "parametric",
          parallel = "multicore",
          .progress = "txt")

# Compile variance and repeatability estimates 
res_varEst <- 
  as_tibble(res_bootOut$t) %>%
  rename("Individual Intercept" = "V1",
         "Series Intercept" = "V2",
         "Series Slope" = "V3",
         "Residual" = "V4",
         "Reaction Norm Intercept Repeatability" = "V5",
         "Long-term Repeatability" = "V6") %>%
  map_df(~ quantile(.x, c((1 - 0.95)/2, 1 - (1 - 0.95)/2),  # calculate 95CI from bootstrap
                  na.rm = TRUE), .id = "scale") %>%
  rename(Variable = scale, Lower95CI = `2.5%`, Upper95CI = `97.5%`) %>%
  mutate(Estimate = pluck(res_bootOut$t0)) %>%  # add R estimates
  relocate(Variable, Estimate) %>%  # reorder columns
  mutate_at(c(2:4), round, 2)  # # round estimates and errors to 2 digits

kable(res_varEst)
```

## Background preference

- Dark background: yes or no

**Create data set with NA values removed**

```{r bkgrd-data}
data_bkgrd_rpt <- 
  remove_dat_na(data_time, c("bkgrd", "t_point_sc", "trial",
                             "sex", "ID", "trial_series", "day",
                             "obsID"))
```

### Random-effect selection

**Test significance of individual intercept (1|ID)**

```{r bkgrd-ID}
# Fit base model
m_bkgrd_base <- 
  glmer(bkgrd ~ t_point_sc * trial + sex +
          (1|ID) + (t_point_sc|trial_series), 
        data = data_bkgrd_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Fit model without individual ID
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|trial_series)")

# LRT for significance bkgrd individual ID
rpt_SigTest(m_bkgrd_base, m_bkgrd_reduc)
```

**Test significance of individual intercept and time slope correlation (t_point_sc|ID)**

- Run before testing random slope because result background slope intercept will determine which slope model to use

```{r bkgrd-ID-time-corr}
# Fit model with intercept-slope correlation
m_bkgrd_full <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series)")  

# Fit model without intercept-slope m_bkgrd_base
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series)")

# LRT for significance bkgrd intercept-slope correlation
rpt_SigTest(m_bkgrd_full, m_bkgrd_reduc)

# print significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_bkgrd_full)$ID,"correlation")[2,1],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2, 
      align = "l")
```

**Test significance of time slope (t_point_sc)**

```{r bkgrd-time}
# Fit model with random slope
m_bkgrd_full <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) +
                                  (t_point_sc|trial_series)") 

# Fit model without random slope
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(1|ID) + (t_point_sc|trial_series)")

# LRT for significance bkgrd random slope 
rpt_SigTest(m_bkgrd_full, m_bkgrd_reduc)
```

**Test significance of (reaction norm) trial series intercept**

```{r bkgrd-rxn-norm-trial-series-int}
# Fit model with random intercept
m_bkgrd_full <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (1|trial_series)")

# Fit model without random intercept
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID)")

# LRT for significance bkgrd intercept
rpt_SigTest(m_bkgrd_full, m_bkgrd_reduc)
```

**Test significance of (reaction norm) trial series intercept and time slope correlation**

- Run before testing random slope b/c result bkgrd slope intercept will determine which slope model to use

```{r bkgrd-rxn-norm-trial-series-time-slope-corr}
# Fit model with random intercept-slope correlation
m_bkgrd_full <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series)")  

# Fit model without random intercept-slope correlation
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (0 + t_point_sc|trial_series) +
                                  (1|trial_series)")

# LRT for significance bkgrd intercept-slope correlation
rpt_SigTest(m_bkgrd_full, m_bkgrd_reduc)

# print significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_bkgrd_full)$trial_series,"correlation")[2,1],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of (reaction norm) time slope**

```{r bkgrd-rxn-norm-time-slope}
# Fit model with random slope
m_bkgrd_full <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series)")  

# Fit model without random slope
m_bkgrd_reduc <- refitModelRanEf(m_bkgrd_base,
                                 "(t_point_sc|ID) + (1|trial_series)")

# LRT for significance bkgrd random slope
rpt_SigTest(m_bkgrd_full, m_bkgrd_reduc)
```

**Final random effect model structure**

- Removed  intercept and time slope correlations for trial series and individual ID due to model convergence errors during bootstrapping

```{r bkgrd-final-rpt-mod}
# Refit without random effect covariances due to singular fits
m_bkgrd_rpt_final <- 
  glmer(bkgrd ~ t_point_sc * trial + sex +
          (0 + t_point_sc|ID) + (1|ID) + (0 + t_point_sc|trial_series) + (1|trial_series),  
        data = data_bkgrd_rpt,
        family = "binomial", 
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))
```

### Variance and repeatability estimates

```{r bkgrd-var-rpt-est, cache=TRUE}
# Bootstrap confidence intervals
res_bootOut <- 
  bootMer(m_bkgrd_rpt_final,  # final rpt model 
          use.u = FALSE, 
          FUN = varEst_bkgrd,  # custom variance estimates function
          seed = 1,
          nsim = 1000,  
          type = "parametric",
          parallel = "multicore",
          .progress = "txt")

# Compile variance and repeatability estimates 
res_varEst <- 
  as_tibble(res_bootOut$t) %>%
  rename("Individual Intercept" = "V1",
         "Individual Slope" = "V2",
         "Series Intercept" = "V3",
         "Series Slope" = "V4",
         "Residual" = "V5",
         "Reaction Norm Intercept Repeatability" = "V6",
         "Reaction Norm Slope Repeatability" = "V7",
         "Long-term Repeatability" = "V8") %>%
  map_df(~ quantile(.x, c((1 - 0.95)/2, 1 - (1 - 0.95)/2),  # calculate 95CI from bootstrap
                  na.rm = TRUE), .id = "scale") %>%
  rename(Variable = scale, Lower95CI = `2.5%`, Upper95CI = `97.5%`) %>%
  mutate(Estimate = pluck(res_bootOut$t0)) %>%  # add R estimates
  relocate(Variable, Estimate) %>%  # reorder columns
  mutate_at(c(2:4), round, 2)  # # round estimates and errors to 2 digits

kable(res_varEst)
```

## Activity

- Number of lines crossed

**Create data set with NA values removed**

```{r act-data}
data_act_rpt <- 
  remove_dat_na(data_act, c("line_cross", "t_point_sc", "trial",
                            "sex", "ID", "trial_series", "day",
                            "obsID"))
```

### Random-effect selection

**Test significance of individual intercept (1|ID)**

```{r act-ID}
# Fit base model
m_act_base <- 
  glmmTMB(line_cross ~ t_point_sc * trial + sex +
            (1|ID) + (t_point_sc|trial_series),
          data = data_act_rpt,
          contrasts = list(trial = c(-1,1), sex = c(-1,1)),
          ziformula= ~ 1,
          family="nbinom2")

# Fit model without individual ID
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(t_point_sc|trial_series)")

# LRT for significance act individual ID
rpt_SigTest(m_act_base, m_act_reduc)
```

**Test significance of individual intercept and time slope correlation (t_point_sc|ID)**

- Run before testing random slope because result act slope intercept will determine which slope model to use

```{r act-ID-time-corr}
# Fit model with intercept-slope correlation
m_act_full <- refitModelRanEf(m_act_base,
                                 "(t_point_sc|ID) + (t_point_sc|trial_series)")  

# Fit model without intercept-slope m_act_base
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(0 + t_point_sc|ID) + (1|ID) +
                                  (t_point_sc|trial_series)")

# LRT for significance act intercept-slope correlation
rpt_SigTest(m_act_full, m_act_reduc)

# print significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_act_full)$cond$ID,"correlation")[1,2],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of time slope (t_point_sc)**

```{r act-time}
# Fit model with random slope
m_act_full <- refitModelRanEf(m_act_base,
                                 "(t_point_sc|ID) + 
                                  (t_point_sc|trial_series)") 

# Fit model without random slope
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (t_point_sc|trial_series)")

# LRT for significance act random slope 
rpt_SigTest(m_act_full, m_act_reduc)

# print non-significant random slope
kable(as_tibble_col(VarCorr(m_act_full)$cond$ID[2,2],
      column_name = "Random slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of (reaction norm) trial series intercept**

```{r act-rxn-norm-trial-series-int}
# Fit model with random intercept
m_act_full <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (1|trial_series)")

# Fit model without random intercept
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(1|ID)")

# LRT for significance act intercept
rpt_SigTest(m_act_full, m_act_reduc)
```

**Test significance of (reaction norm) trial series intercept and time slope correlation**

- Run before testing random slope b/c result act slope intercept will determine which slope model to use

```{r act-rxn-norm-trial-series-time-slope-corr}
# Fit model with random intercept-slope correlation
m_act_full <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (t_point_sc|trial_series)")  

# Fit model without random intercept-slope correlation
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (0 + t_point_sc|trial_series) +
                                  (1|trial_series)")

# LRT for significance act intercept-slope correlation
rpt_SigTest(m_act_full, m_act_reduc)

# print significant intercept-slope correlation
kable(as_tibble_col(attr(VarCorr(m_act_full)$cond$ID,"correlation")[1,2],
      column_name = "Random intercept-slope variance estimate"),
      digits = 2,
      align = "l")
```

**Test significance of (reaction norm) time slope**

```{r act-rxn-norm-time-slope}
# Fit model with random slope
m_act_full <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (t_point_sc|trial_series)")  

# Fit model without random slope
m_act_reduc <- refitModelRanEf(m_act_base,
                                 "(1|ID) + (1|trial_series)")

# LRT for significance act random slope
rpt_SigTest(m_act_full, m_act_reduc)
```

**Final random effect model structure**

- Removed trial series intercept and time slope correlation due to model convergence errors during bootstrapping

```{r act-final-rpt-mod}
m_act_rpt_final <- 
  glmmTMB(line_cross ~ t_point_sc * trial + sex +
            (1|ID) + (0 + t_point_sc|trial_series) + (1|trial_series),
          data = data_act_rpt,
          contrasts = list(trial = c(-1,1), sex = c(-1,1)),
          ziformula= ~ 1,
          family="nbinom2")
```

### Variance and repeatability estimates

```{r act-var-rpt-est, cache=TRUE}
# Bootstrap confidence intervals
res_bootOut <- 
  bootMer(m_act_rpt_final,  # final rpt model 
          use.u = FALSE, 
          FUN = varEst_act,  # custom variance estimates function
          seed = 1,
          nsim = 1000,  
          type = "parametric",
          parallel = "multicore",
          .progress = "txt")

# Compile variance and repeatability estimates 
res_varEst <- 
  as_tibble(res_bootOut$t) %>%
  rename("Individual Intercept" = "V1",
         "Series Intercept" = "V2",
         "Series Slope" = "V3",
         "Residual" = "V4",
         "Reaction Norm Intercept Repeatability" = "V5",
         "Long-term Repeatability" = "V6") %>%
  map_df(~ quantile(.x, c((1 - 0.95)/2, 1 - (1 - 0.95)/2),  # calculate 95CI from bootstrap
                  na.rm = TRUE), .id = "scale") %>%
  rename(Variable = scale, Lower95CI = `2.5%`, Upper95CI = `97.5%`) %>%
  mutate(Estimate = pluck(res_bootOut$t0)) %>%  # add R estimates
  relocate(Variable, Estimate) %>%  # reorder columns
  mutate_at(c(2:4), round, 2)  # # round estimates and errors to 2 digits

kable(res_varEst)
```