# Analyses: Predictors of behavior 
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Analyses for predictors of behavior for Lafferty et al. 2020

```{r pred-behav-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE, cache = TRUE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE, cache = TRUE)
```

## Setup

```{r pred-behav-setup}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(rptR)  # for repeatability analyses
library(lme4)
library(lmerTest)
library(DHARMa)
library(broom.mixed)  # for tidying model output
library(knitr)
library(glmmTMB)  # for activity analyses

# Load personalized functions
source("R/custom-functions.R")

# Read in data 
# Data collected once each day
data_day <-
  read_csv("data/data_day_LaffertyEtAl_2020.csv",
           col_types = list(exit_fast = col_factor(),
                            trial = col_factor(levels = c("1", "2")),
                            sex = col_factor(levels = c("Female", "Male")))) %>%
  mutate(obsID = 1:nrow(.))  # observation level random effect

# Data collected at time points throughout the day
data_time <-
  read_csv("data/data_time_LaffertyEtAl_2020.csv",
             col_types = list(trial = col_factor(levels = c("1", "2")),
                              sex = col_factor(levels = c("Female", "Male")),
                              of_loc = col_factor(levels = c("Inner", "Outer")),
                              bkgrd = col_factor(levels = c("Light", "Dark")))) %>%
  mutate(obsID = 1:nrow(.),  # observation level random effect
         across(where(is.character), ~ as_factor(.x)))  

# Data collected at time points throughout the day for lines crossed
data_act <-
  read_csv("data/data_activity_LaffertyEtAl_2020.csv",
             col_types = list(trial = col_factor(levels = c("1", "2")),
                              sex = col_factor(levels = c("Female", "Male")),
                              loc = col_factor(levels = c("Inner", "Outer")),
                              backg = col_factor(levels = c("Light", "Dark")))) %>%
  mutate(obsID = 1:nrow(.),  # observation level random effect
         across(where(is.character), ~ as_factor(.x)))  
```

### Area explored

- Proportion of unique tiles entered

**Create dataset without NAs**

```{r pred-behav-expl-dat}
data_expl <- 
  remove_dat_na(data_day, c("uniq_tiles", "tile_fails", "trial",
                            "sex", "ID", "day", "obsID", "mass_sc")) %>%
  mutate(expl.y = cbind(uniq_tiles, tile_fails))
```

**Predictor interaction selection**

```{r explore-predict}
# Fit model
m_expl_full <- 
  glmer(expl.y ~ trial + sex * mass_sc + (1|ID),
        family = "binomial", 
        data = data_expl,
        contrasts = list(trial = c(-1,1)),
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Test for significance of interactions
drop1(m_expl_full, test = "Chisq") %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Drop sex * mass interaction
m_expl_final <- update(m_expl_full, ~ . -sex:mass_sc) 

# Tidy model output
broom.mixed::tidy(m_expl_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (exp)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of area explored")

# R squared (delta)
kable(t(MuMIn::r.squaredGLMM(m_expl_final)[2,]), digits = 2, align = "l",
      caption = "Marginal and conditional R^2") 
```

**Grand mean**

```{r explore-mean}
# Fit model
m_expl_means <- 
  glmer(expl.y ~ trial + sex + mass_sc + (1|ID),
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),  # Change contrasts for sex and trial
        family = "binomial", 
        data = data_expl,
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Tidy model output
broom.mixed::tidy(m_expl_means,
                  effects = "fixed") %>%
  mutate(estimate.pr = round_est(plogis(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.pr, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (proportion)` = estimate, -estimate.pr) %>%
  filter(term == "(Intercept)") %>%
  kable(., align = "l",
        caption = "Marginal mean area explored")
```

### Risk-taking

#### Shelter exit latency

- Exit immediately: yes or no

**Create data set with NA values removed**

```{r shel-exit-dat}
data_shelExit <- 
  remove_dat_na(data_day, c("exit_fast", "trial",
                            "sex", "ID", "day",
                            "obsID", "mass_sc"))
```

**Predictor interaction selection**

```{r shel-exit-predict}
# Fit model
m_ExitLat_full <- 
  glmer(exit_fast ~ trial + sex * mass_sc + (1|ID),
        family = "binomial", 
        data = data_shelExit,
        contrasts = list(trial = c(-1,1)),
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Test for significance of interactions
drop1(m_ExitLat_full, test = "Chisq") %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)
  
# Drop sex * mass interaction
m_ExitLat_final <- update(m_ExitLat_full, ~ . -sex:mass_sc) 

# Tidy model output
broom.mixed::tidy(m_ExitLat_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (exp)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of shelter exit latency")

# R squared (delta)
kable(t(MuMIn::r.squaredGLMM(m_ExitLat_final)[2,]), digits = 2, align = "l",
      caption = "Marginal and conditional R^2")
```

**Grand mean**

```{r shel-exit-mean}
# Fit model
m_ExitLat_means <- 
  glmer(exit_fast ~ trial + sex + mass_sc + (1|ID),
        contrasts = list(trial = c(-1,1), sex = c(-1,1)),  # Change contrasts for sex and trial
        family = "binomial", 
        data = data_shelExit,
        glmerControl(optimizer="bobyqa",
        optCtrl = list(maxfun = 1e6)))

# Tidy model output
broom.mixed::tidy(m_ExitLat_means,
                  effects = "fixed") %>%
  mutate(estimate.pr = round_est(plogis(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.pr, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (probability)` = estimate, -estimate.pr) %>%
  filter(term == "(Intercept)") %>%
  kable(., align = "l",
        caption = "Marginal mean shelter exit latency")
```

#### Shelter preference

- Inside or outside the shelter

**Create data set with NA values removed**

```{r shel-pref-data}
data_shel <- 
  remove_dat_na(data_time, c("shelter", "t_point_sc", "trial", "mass_sc",
                             "sex", "ID", "trial_series", "day", "bkgrd",
                             "obsID"))
```

**Predictor interaction selection**

```{r shel-pref-predict, cache=TRUE}
# Fit full model
m_shel_base <- 
  glmer(shelter ~ t_point_sc * trial + 
          sex * mass_sc +
          sex * t_point_sc +
          bkgrd * t_point_sc +
          sex * bkgrd +
          (t_point_sc|trial_series),  
        data = data_shel,
        family = "binomial", 
        contrasts = list(trial = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Find largest non-significant p-value for interaction
drop1(m_shel_base, test = "Chisq") %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Remove largest non-significant p-value for interaction and update model and continue process
m1_shel_tmp <- update(m_shel_base, ~ . -t_point_sc:trial)  # Remove most non-sig interactions
m1_shel_LRT <- drop1(m1_shel_tmp, test = "Chisq")  # Update model and check remaining sig interactions
m1_shel_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Same as above
m2_shel_tmp <- update(m1_shel_tmp, ~ . -sex:bkgrd)  # same as above
m2_shel_LRT <- drop1(m2_shel_tmp, test = "Chi")  # same as above
m2_shel_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Same as above
m3_shel_tmp <- update(m2_shel_tmp, ~ . -t_point_sc:bkgrd)  # same as above
m3_shel_LRT <- drop1(m3_shel_tmp, test = "Chi")  # same as above
m3_shel_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Same as above
m4_shel_tmp <- update(m3_shel_tmp, ~ . -t_point_sc:sex)  # same as above
m4_shel_LRT <- drop1(m4_shel_tmp, test = "Chi")  # same as above
m4_shel_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Same as above
m5_shel_tmp <- update(m4_shel_tmp, ~ . -sex:mass_sc)  # same as above
m5_shel_LRT <- drop1(m5_shel_tmp, test = "Chi")  # same as above
m5_shel_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Fit final model of predictors
m_shel_final <- update(m_shel_base, formula(m5_shel_tmp))

# Tidy model output
broom.mixed::tidy(m_shel_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (odds ratio)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of shelter preference")

# R squared (delta)
kable(t(MuMIn::r.squaredGLMM(m_shel_final)[2,]), digits = 2, align = "l",
      caption = "Marginal and conditional R^2")
```

##### Grand Mean

```{r shel-pref-marg}
m_shel_final_marg <- update(m_shel_final,
                            contrasts = list(trial = c(-1,1), sex = c(-1,1),
                                             bkgrd = c(-1,1)))

m_shel_final_marg %>%
  pretty_PredictTab(., mixedModel = TRUE, kable = FALSE) %>%
  mutate(estimate = paste(round(exp(as.numeric(estimate)), digits = 3),
                          " (",
                          round(plogis(as.numeric(estimate)), digits = 3),
                          ")", sep = "")) %>%
  filter(term == "(Intercept)") %>%
  select(-effect, `estimate odds ratio (probability)` = estimate) %>%
  kable(., align = "l")
```

#### Open-Field location preference

- Inner or Outer area of the Open-Field

**Create data set with NA values removed**

```{r OF-pref-data}
data_OF <- 
  remove_dat_na(data_time, c("of_loc", "t_point_sc", "trial", "mass_sc",
                             "sex", "ID", "trial_series", "day", "bkgrd",
                             "obsID"))
```

**Predictor interaction selection**

```{r OF-pref-predict, cache=TRUE}
# Fit full model
m_OF_base <- 
  glmer(of_loc ~ t_point_sc * trial + 
          sex * mass_sc +
          sex * t_point_sc +
          bkgrd * t_point_sc +
          sex * bkgrd +
          (1|ID) + (t_point_sc|trial_series),  
        data = data_OF,
        family = "binomial", 
        contrasts = list(trial = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Find largest non-significant p-value for interaction
drop1(m_OF_base, test = "Chisq") %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Remove largest non-significant p-value for interaction and update model and continue process
m1_OF_tmp <- update(m_OF_base, ~ . -t_point_sc:sex)  # Remove most non-sig interactions
m1_OF_LRT <- drop1(m1_OF_tmp, test = "Chisq")  # Update model and check remaining sig interactions
m1_OF_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Same as above
m2_OF_tmp <- update(m1_OF_tmp, ~ . -sex:mass_sc)  # same as above
m2_OF_LRT <- drop1(m2_OF_tmp, test = "Chi")  # same as above
m2_OF_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Same as above
m3_OF_tmp <- update(m2_OF_tmp, ~ . -t_point_sc:trial)  # same as above
m3_OF_LRT <- drop1(m3_OF_tmp, test = "Chi")  # same as above
m3_OF_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Fit final model of predictors
m_OF_final <- update(m_OF_base, formula(m3_OF_tmp))

# Tidy model output
broom.mixed::tidy(m_OF_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (odds ratio)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of Open-Field location preference")

# R squared (delta)
kable(t(MuMIn::r.squaredGLMM(m_OF_final)[2,]), digits = 2, align = "l",
      caption = "Marginal and conditional R^2")
```

### Background preference preference

- Light or dark background

**Create data set with NA values removed**

```{r bkgrd-pref-data}
data_bkgrd <- 
  remove_dat_na(data_time, c("bkgrd", "t_point_sc", "trial", "mass_sc",
                             "sex", "ID", "trial_series", "day", "obsID"))
```

**Predictor interaction selection**

```{r bkgrd-pref-predict, cache=TRUE}
# Fit full model
m_bkgrd_base <- 
  glmer(bkgrd ~ t_point_sc * trial + 
          sex * mass_sc +
          sex * t_point_sc +
          (1|ID) + (0 + t_point_sc|ID) + (1|trial_series) + (0 + t_point_sc|trial_series),
        data = data_bkgrd,
        family = "binomial", 
        contrasts = list(trial = c(-1,1)),
        glmerControl(optimizer="bobyqa",
                    optCtrl = list(maxfun = 1e6)))

# Find largest non-significant p-value for interaction
drop1(m_bkgrd_base, test = "Chisq") %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Remove largest non-significant p-value for interaction and update model and continue process
m1_bkgrd_tmp <- update(m_bkgrd_base, ~ . -t_point_sc:trial)  # Remove most non-sig interactions
m1_bkgrd_LRT <- drop1(m1_bkgrd_tmp, test = "Chisq")  # Update model and check remaining sig interactions
m1_bkgrd_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  kable_title(.)

# Same as above
m2_bkgrd_tmp <- update(m1_bkgrd_tmp, ~ . -t_point_sc:sex)  # same as above
m2_bkgrd_LRT <- drop1(m2_bkgrd_tmp, test = "Chi")  # same as above
m2_bkgrd_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Same as above
m3_bkgrd_tmp <- update(m2_bkgrd_tmp, ~ . -sex:mass_sc)  # same as above
m3_bkgrd_LRT <- drop1(m3_bkgrd_tmp, test = "Chi")  # same as above
m3_bkgrd_LRT %>%
  rd_stepwise_out(., glm = TRUE) %>%
  knitr::kable(.)

# Fit final model bkgrd predictors
m_bkgrd_final <- update(m_bkgrd_base, formula(m3_bkgrd_tmp))

# Tidy model output
broom.mixed::tidy(m_bkgrd_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  select(-effect, `estimate (odds ratio)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of background preference")

# R squared (delta)
kable(t(MuMIn::r.squaredGLMM(m_bkgrd_final)[2,]), digits = 2, align = "l",
      caption = "Marginal and conditional R^2")
```

### Activity

- Number of lines crossed

**Create data set with NA values removed**

```{r act-pref-data}
data_act <- 
  remove_dat_na(data_act, c("line_cross", "t_point_sc", "trial", "mass_sc",
                             "sex", "ID", "trial_series", "day", "obsID",
                             "backg", "loc", "shel_dur"))
```

**Predictor interaction selection**

```{r act-pref-predict, cache=TRUE}
# Fit full model
m_act_base <- 
  glmmTMB(line_cross ~ t_point_sc * trial + 
            sex * mass_sc +
            sex * t_point_sc +
            sex * backg +
            sex * loc +
            backg * t_point_sc +
            loc * backg +
            loc * t_point_sc +
            (1|ID) + (t_point_sc|trial_series),
          data = data_act,
          ziformula= ~ 1,
          family="nbinom2", 
          contrasts = list(trial = c(-1,1)))

# Find largest non-significant p-value for interaction
m_act_LRT <- 
  drop1(m_act_base, test = "Chisq") %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  kable_title(.)

m_act_LRT

# Remove largest non-significant p-value for interaction and update model and continue process
m1_act_tmp <- update(m_act_base, ~ . -backg:loc)  # Remove most non-sig interactions
m1_act_LRT <- drop1(m1_act_tmp, test = "Chisq")  # Update model and check remaining sig interactions
m1_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  kable_title(.)

# Same as above
m2_act_tmp <- update(m1_act_tmp, ~ . -t_point_sc:backg)  # same as above
m2_act_LRT <- drop1(m2_act_tmp, test = "Chi")  # same as above
m2_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Same as above
m3_act_tmp <- update(m2_act_tmp, ~ . -t_point_sc:trial)  # same as above
m3_act_LRT <- drop1(m3_act_tmp, test = "Chi")  # same as above
m3_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Same as above
m4_act_tmp <- update(m3_act_tmp, ~ . -t_point_sc:sex)  # same as above
m4_act_LRT <- drop1(m4_act_tmp, test = "Chi")  # same as above
m4_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Same as above
m5_act_tmp <- update(m4_act_tmp, ~ . -sex:backg)  # same as above
m5_act_LRT <- drop1(m5_act_tmp, test = "Chi")  # same as above
m5_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Same as above
m6_act_tmp <- update(m5_act_tmp, ~ . -t_point_sc:loc)  # same as above
m6_act_LRT <- drop1(m6_act_tmp, test = "Chi")  # same as above
m6_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Same as above
m7_act_tmp <- update(m6_act_tmp, ~ . -sex:mass_sc)  # same as above
m7_act_LRT <- drop1(m7_act_tmp, test = "Chi")  # same as above
m7_act_LRT %>%
  rd_stepwise_out(., glmmTMB = TRUE) %>%
  knitr::kable(.)

# Fit final model act predictors
m_act_final <- update(m_act_base, formula(m7_act_tmp))

# Tidy model output
broom.mixed::tidy(m_act_final,
                  effects = "fixed") %>%
  mutate(estimate.exp = round_est(exp(estimate)),
         estimate = paste(round_est(estimate)," (",estimate.exp, ")", sep = ""),
         std.error = round_est(std.error),
         statistic = round_est(statistic),
         p.value = round_pval(p.value)) %>%
  filter(component == "cond") %>%
  select(-effect, -component, `estimate (count rate)` = estimate, -estimate.exp) %>%
  kable(., align = "l",
        caption = "Predictors of activity")
```

- R^2^ unavailable for glmmTMB